---
title: "Case Study 2"
author: "Samuel Vonpaays Soh"
date: "2/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(caret)
library(e1071)
library(class)
library(skimr)
library(kableExtra)

url <- "https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%2014%20and%2015%20Case%20Study%202/CaseStudy2-data.csv"
url1 <- "https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%2014%20and%2015%20Case%20Study%202/CaseStudy2CompSet%20No%20Attrition.csv"

casestudy2 <- read.csv(url, stringsAsFactors = TRUE, header = TRUE)
casestudy2_No_Attrition <- read.csv(url1, stringsAsFactors = TRUE, header = TRUE)
casestudy2_No_Salary <- read.csv("/Users/samuelvonpaayssoh/Desktop/NoSalary.csv", stringsAsFactors = TRUE, header = TRUE)
dim(casestudy2)
dim(casestudy2_No_Attrition)
dim(casestudy2_No_Salary)
```
## 870 rows and 36 columns

## EDA
```{r }
colnames(casestudy2)
colnames(casestudy2_No_Attrition)
colnames(casestudy2_No_Salary)
str(casestudy2)
summary(casestudy2)
summary(casestudy2_No_Attrition)
summary(casestudy2_No_Salary)
```

## Checking N/A or missing values

```{r }
sum(is.na(casestudy2))
sum(is.na(casestudy2_No_Attrition))
sum(is.na(casestudy2_No_Salary))
```
## Removing dummy columns:non valuable columns
```{r }
casestudy2<- casestudy2 %>% select(-one_of(c("EmployeeCount","EmployeeNumber","StandardHours","Over18")))
No_Attrition <- casestudy2_No_Attrition %>% select(-one_of(c("EmployeeCount","EmployeeNumber","StandardHours","Over18")))
No_Salary <- casestudy2_No_Salary %>% select(-one_of(c("EmployeeCount","EmployeeNumber","StandardHours","Over18")))
colnames(casestudy2)
dim(casestudy2)
str(casestudy2)
str(No_Attrition)
str(No_Salary)
```
## Converting int to factors
```{r }
casestudy2$Education <- as.factor(casestudy2$Education)
casestudy2$EnvironmentSatisfaction <- as.factor(casestudy2$EnvironmentSatisfaction)
casestudy2$JobInvolvement <- as.factor(casestudy2$JobInvolvement)
casestudy2$JobLevel <- as.factor(casestudy2$JobLevel)
casestudy2$JobSatisfaction <- as.factor(casestudy2$JobSatisfaction)
casestudy2$PerformanceRating <- as.factor(casestudy2$PerformanceRating)
casestudy2$RelationshipSatisfaction <- as.factor(casestudy2$RelationshipSatisfaction)
casestudy2$StockOptionLevel <- as.factor(casestudy2$StockOptionLevel)
casestudy2$TrainingTimesLastYear <- as.factor(casestudy2$TrainingTimesLastYear)
casestudy2$WorkLifeBalance <- as.factor(casestudy2$WorkLifeBalance)

No_Attrition$Education <- as.factor(No_Attrition$Education)
No_Attrition$EnvironmentSatisfaction <- as.factor(No_Attrition$EnvironmentSatisfaction)
No_Attrition$JobInvolvement <- as.factor(No_Attrition$JobInvolvement)
No_Attrition$JobLevel <- as.factor(No_Attrition$JobLevel)
No_Attrition$JobSatisfaction <- as.factor(No_Attrition$JobSatisfaction)
No_Attrition$PerformanceRating <- as.factor(No_Attrition$PerformanceRating)
No_Attrition$RelationshipSatisfaction <- as.factor(No_Attrition$RelationshipSatisfaction)
No_Attrition$StockOptionLevel <- as.factor(No_Attrition$StockOptionLevel)
No_Attrition$TrainingTimesLastYear <- as.factor(No_Attrition$TrainingTimesLastYear)
No_Attrition$WorkLifeBalance <- as.factor(No_Attrition$WorkLifeBalance)

No_Salary$Education <- as.factor(No_Salary$Education)
No_Salary$EnvironmentSatisfaction <- as.factor(No_Salary$EnvironmentSatisfaction)
No_Salary$JobInvolvement <- as.factor(No_Salary$JobInvolvement)
No_Salary$JobLevel <- as.factor(No_Salary$JobLevel)
No_Salary$JobSatisfaction <- as.factor(No_Salary$JobSatisfaction)
No_Salary$PerformanceRating <- as.factor(No_Salary$PerformanceRating)
No_Salary$RelationshipSatisfaction <- as.factor(No_Salary$RelationshipSatisfaction)
No_Salary$StockOptionLevel <- as.factor(No_Salary$StockOptionLevel)
No_Salary$TrainingTimesLastYear <- as.factor(No_Salary$TrainingTimesLastYear)
No_Salary$WorkLifeBalance <- as.factor(No_Salary$WorkLifeBalance)
```

```{r}
str(casestudy2)
str(No_Attrition)
str(No_Salary)
```

```{r}
casestudy2_numeric <- casestudy2 %>% select_if(is.numeric)
colnames(casestudy2_numeric)
```
##Department Turnover
```{r , include=TRUE}
Attrition_Department <- casestudy2 %>% count(Department,Attrition)
spread(Attrition_Department,Attrition,n) %>% mutate(sum=No+Yes) %>% mutate(attrition_pct=Yes/sum) %>% ggplot(aes(x=Department,y=attrition_pct)) +geom_col()
```
## Age Turnover
```{r , include=TRUE}
Attrition_by_Age <- casestudy2 %>% count(Age,Attrition)
spread(Attrition_by_Age,Attrition,n) %>% mutate(sum=No+Yes) %>% mutate(attrition_pct=Yes/sum) %>% ggplot(aes(x=Age , y=attrition_pct)) +geom_point() +geom_smooth(method="loess") +ylab("Attrition")
```
## Job Satisfaction Turnover
```{r , include=TRUE}
Attrition_by_JobSatisfaction <- casestudy2 %>% count(JobSatisfaction,Attrition)
spread(Attrition_by_JobSatisfaction,Attrition,n) %>% mutate(sum=No+Yes) %>% mutate(attrition_pct=Yes/sum) %>% ggplot(aes(x=JobSatisfaction , y=attrition_pct)) +geom_point() + ylab("Attrition")
```

## Job Role Turnover
```{r , include=TRUE}
Attrition_by_JobRole <- casestudy2 %>% count(JobRole,Attrition)
spread(Attrition_by_JobRole,Attrition,n) %>% mutate(sum=No+Yes) %>% mutate(attrition_pct=Yes/sum) %>% ggplot(aes(x=reorder(JobRole,-attrition_pct) , y=attrition_pct)) +geom_col() + coord_flip()+ ylab("Attrition")+xlab("Job Role")
```
## Working Years Turnover
```{r , include=TRUE}
Attrition_by_totalworkingyears <- casestudy2 %>% count(TotalWorkingYears,Attrition)
spread(Attrition_by_totalworkingyears,Attrition,n) %>% mutate(sum=No+Yes) %>% mutate(attrition_pct=Yes/sum) %>% ggplot(aes(x=TotalWorkingYears , y=attrition_pct)) +geom_point()+ ylab("Attrition")
```
## Increased Salary Turnover
```{r , include=TRUE}
Attrition_by_salaryincreased <- casestudy2 %>% count(PercentSalaryHike,Attrition)
spread(Attrition_by_salaryincreased,Attrition,n) %>% mutate(sum=No+Yes) %>% mutate(attrition_pct=Yes/sum) %>% ggplot(aes(x=PercentSalaryHike , y=attrition_pct)) +geom_point()+ ylab("Attrition")
```
## Hourly Rate Turnover
```{r , include=TRUE}
Attrition_by_hourly <- casestudy2 %>% count(HourlyRate,Attrition)
spread(Attrition_by_hourly,Attrition,n) %>% mutate(sum=No+Yes) %>% mutate(attrition_pct=Yes/sum) %>% ggplot(aes(x=HourlyRate , y=attrition_pct)) +geom_point()+ ylab("Attrition")
```
## Monthly Income Turnover
```{r , include=TRUE}
Attrition_by_income <- na.omit(casestudy2) %>% count(MonthlyIncome,Attrition)
spread(Attrition_by_income,Attrition,n) %>% mutate(sum=No+Yes) %>% mutate(attrition_pct=Yes/sum) %>% ggplot(aes(x=MonthlyIncome ,y=attrition_pct)) +geom_point()+ ylab("Attrition")
```
## Overtime Turnover
```{r , include=TRUE}
Attrition_by_overtime <- casestudy2 %>% count(OverTime,Attrition)
spread(Attrition_by_overtime,Attrition,n) %>% mutate(sum=No+Yes) %>% mutate(attrition_pct=Yes/sum) %>% ggplot(aes(x=OverTime , y=attrition_pct)) +geom_col()+ ylab("Attrition")
```


##Naive Bayes model on Age,Job Role, Job Satisfaction, Over Time
```{r , include=TRUE}

Accuracy_Holder =numeric(100)
Sensitivity_Holder =numeric(100)
Specificity_Holder = numeric(100)

for(i in 1: 100)
  {
set.seed(i)
trainIndices=sample(seq(1:length(casestudy2$Age)),round(.85*length(casestudy2$Age)))
trainNB=casestudy2[trainIndices,]
testNB=casestudy2[-trainIndices,]
model=naiveBayes(trainNB[,c("Age","JobRole","JobSatisfaction","OverTime")],factor(trainNB$Attrition,labels = c("No","Yes")))
CM_NB = confusionMatrix(table(factor(testNB$Attrition, labels = c("No", "Yes")),predict(model,testNB[,c("Age","JobRole","JobSatisfaction","OverTime")])))
CM_NB

Accuracy_Holder[i] = CM_NB$overall[1]
Sensitivity_Holder[i] = CM_NB$byClass[1]
Specificity_Holder[i] = CM_NB$byClass[2]
}

mean(Accuracy_Holder)
#Standard Error of the Mean
#sd(AccHolder)/sqrt(100) 
mean(Sensitivity_Holder)
mean(Specificity_Holder)

```
##KNN
```{r}
AccHolder = numeric(100)
SensHolder = numeric(100)
SpecHolder = numeric(100)
csClean <- na.omit(casestudy2)
splitPerc = .60
iterations = 100
numks = 60
masterAcc = matrix(nrow = iterations, ncol = numks)
  
for(j in 1:iterations)
{
accs = data.frame(accuracy = numeric(60), k = numeric(60))
  #set.seed(i)

trainIndices = sample(1:dim(csClean)[1],round(splitPerc * dim(csClean)[1]))
train = csClean[trainIndices,]
test = csClean[-trainIndices,]
for(i in 1:numks)
{
classifications = knn(train[,c("JobSatisfaction","Age")],test[,c("JobSatisfaction","Age")],as.factor(train$Attrition), prob = TRUE, k = i)
table(as.factor(test$Attrition),classifications)
CM = confusionMatrix(table(as.factor(test$Attrition),classifications))
masterAcc[j,i]=CM$overall[1]
}
}

MeanAcc = colMeans(masterAcc)
plot(seq(1,numks,1),MeanAcc,type="l")
which.max(MeanAcc)
max(MeanAcc)
#AccHolder = numeric(100)
#SensHolder = numeric(100)
#SpecHolder = numeric(100)

for(i in 1:iterations){
  set.seed(i)
  trainIndices = sample(1:dim(csClean)[1],round(splitPerc * dim(csClean)[1]))
  train = csClean[trainIndices,]
  test = csClean[-trainIndices,]
  classifications = knn(train[,c("JobSatisfaction","Age")],test[,c("JobSatisfaction","Age")],as.factor(train$Attrition), prob = TRUE, k = 7)
  table(as.factor(test$Attrition),classifications)
  CM_KNN = confusionMatrix(table(as.factor(test$Attrition),classifications))
  AccHolder[i] = CM_KNN$overall[1]
  SensHolder[i] = CM_KNN$byClass[1]
  SpecHolder[i] = CM_KNN$byClass[2]
}
mean(AccHolder)
mean(SensHolder)
mean(SpecHolder)
```

```{r}
model=naiveBayes(casestudy2[,c("Age","JobRole","JobSatisfaction","OverTime")],factor(casestudy2$Attrition,labels = c("No","Yes")))
No_Attrition$Attrition=predict(model,No_Attrition[,c("Age","JobRole","JobSatisfaction","OverTime")])
Output1 <- No_Attrition 
write.csv(Output1,file = "/Users/samuelvonpaayssoh/Desktop/Case2PredictionVonPaaysSoh Attrition.csv", row.names=FALSE)
```


```{r}

library(GGally)
Salaryvars <- c("MonthlyIncome","Age", "Gender", "JobLevel", "JobRole", "BusinessTravel","StockOptionLevel", "TotalWorkingYears")
casestudy2_Salary <- casestudy2[Salaryvars]
casestudy2_Salary %>% ggpairs()
```

```{r}
trainIndices_salary=sample(1:dim(casestudy2)[1],round(.7*dim(casestudy2)[1]))
train_salary= casestudy2[trainIndices_salary,]
test_salary=casestudy2[-trainIndices_salary,]
fit=lm(MonthlyIncome ~ JobLevel+TotalWorkingYears, data=train_salary)
summary(fit)

Prediction <- predict(fit, test_salary)
RMSE(test_salary$MonthlyIncome, Prediction)
Prediction_Salary <- predict(fit,No_Salary)
Salary_Output = cbind.data.frame(No_Salary$ID,Prediction_Salary)
colnames(Salary_Output) <= c("ID","MonthlyIncome")
write.csv(Salary_Output, file ="/Users/samuelvonpaayssoh/Desktop/Case2PredictionVonPaaysSoh Salary.csv", row.names=FALSE )
  
  
```

```{r}

```